\makeindex[name=spells,title={Spell Summaries},columns=2]

% for CS callout boxes

\newcommand\commentary[3]{
  \node[text=white, overlay, rectangle callout, callout relative pointer={(#2)}, fill=black,] (randomNode) at (#1) [text width=4cm]{#3};
}

\newcommand{\racechart}{

\begin{nametable}[clX]{Race}
  
  \textbf{Roll} & \textbf{Race} & \textbf{Adjustments} \\\hline

  2-3 & Gnome & \mbox{+1 Intelligence, +1 Dexterity,} \mbox{Strength -2, Speed -1} \\

  4-5 & Dwarf & +1 Dexterity, -1 Speed \\

  6-8 & Human & +1 Strength, -1 Wits \\

  9-10 & Elf & +1 Wits, -1 Strength \\

  11-12 & Gnoll & \mbox{+1 Strength, +1 Speed,} \mbox{-1 Intelligence, -2 Charisma} \\

\end{nametable}
}

\newcommand\attributeChart{
  \begin{boxtable}[YcL]

    \textbf{Result} & \textbf{Bonus} & \textbf{Meaning} \\
    \hline

    2 & -3 & Pathetic \\

    3 & -2 & Useless \\

    4-5 & -1 & Weak \\

    6-8 & 0 & Normal \\

    9-10 & +1 & Exceptional \\

    11 & +2 & Excellent \\

    12 & +3 & Peak \\

  \end{boxtable}
}

\newcommand\XPchart{
  \begin{nametable}[XYYYYY,float*=b]{\Gls{xp} Chart}
    \textbf{Level} &  \textbf{Attributes}  &  \textbf{Skill}  & \textbf{Martial Skills}  & \textbf{Knacks} & \textbf{Sphere}  \\
    \hline
    <1                &   5        &         &                      &                                & \\
    1st               &   10       &   5     &        10            &                             5  &    10  \\
    2nd               &   20       &   10    &        20            &                            10  &    15  \\
    3rd               &   30       &   15    &        40            &                            15  &    25  \\
    4th               &   50       &         &                      &                            20  &    45  \\
    5th               &            &         &                      &                            25  &    85  \\

  \end{nametable}
}

\renewcommand\csComments{
    \commentary{[xshift=12,yshift=-9em]TCBPOSTER@title.north}{-2em,2em}{Write down your name and roll a random race (\autopageref{raceRoll}).}

    \commentary{[xshift=4em,yshift=-2em]TCBPOSTER@attributes.south east}{-4.1em,2.8em}{Roll to determine each of your Attributes, then apply the racial modifier. The large `\statDot' represents a Bonus of `0'.}

    \commentary{[xshift=4em,yshift=-1em]TCBPOSTER@gumption.west}{-20em,9em}{}
    \commentary{[xshift=4em,yshift=0em]TCBPOSTER@gumption.west}{1em,-28em}{}
    \commentary{[xshift=4em,yshift=0em]TCBPOSTER@gumption.west}{-4em,6.5em}{Spend 50 \glspl{xp} to purchase Attributes, Skills, and Knacks (\autopageref{xp}).}

    \commentary{[xshift=3em,yshift=-1em]TCBPOSTER@gumption.south}{.4em,5em}{\small Place coins in the circles to keep track of the current score.}

    \commentary{[xshift=3em,yshift=2em]TCBPOSTER@derived.south}{3em,2em}{\small Place a coin on the \glspl{ap} tracker to keep track of your \glspl{ap} during combat.}

    \commentary{[xshift=-7em,yshift=-6em]TCBPOSTER@title.east}{-2em,1em}{Write down your Code and Culture.}

    \commentary{[xshift=5em,yshift=5em]TCBPOSTER@armoury.north west}{4em,-25em}{}
    \commentary{[xshift=5em,yshift=5em]TCBPOSTER@armoury.north west}{-1em,-2.5em}{Put your weapons here, and check their stats.}
}


\newcommand\acquisitionXP{

  \begin{xpchart}{Acquisition}

    1 & Buying an expensive item -- worth 1 \gls{gp} or more. \\

    1 & Being greeted deferentially by a stranger. \\

    1 & Confirming you are the richest person in a new region. \\

    1 & Hiring a new servant. \\

    1 & Meeting a famous person. \\

    3 & Buying a very expensive item -- at least 10 \gls{gp} in value. \\

    3 & Disarming a potential conflict. \\

    5 & Starting a new business. \\

    5 & Finding a magical item. \\

    5 & Gaining a new title such as a guild master or townmaster. \\

    5 & Buying an extremely expensive item -- at least 20 \gls{gp} in value. \\

    10 & Marrying into a prestigious family. \\

    10 & Acquiring a horde of new wealth -- at least 100\% of the character's maximum lifetime total, and at least 1 gp. \\

    10 & Buying an expensive new home -- at least 200 \gls{gp} in value. \\

  \end{xpchart}
}

\newcommand\justiceXP{
  \begin{xpchart}{\Glsentrytext{justicegod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Enforcing a law. \\

    1 & Feeding the poor. \\

    1 & Hosting guests. \\

    1 & Punishing law breakers. \\

    1 & Returning someone's valuables to them. \\

    1 & Enduring a troubling oath. \\

    3 & Enforcing a major law or imposing the law on a group. \\

    3 & Donating at least 10 \gls{gp} to the temple. \\

    3 & Returning from a dangerous mission with all allies unharmed. \\

    5 & Completing a quest you swore to complete over one session ago. \\

    5 & Correcting the local law. \\

    5 & Creating a peace treaty between factions in danger of fighting. \\

    10 & Creating a peace treaty between warring factions. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    15 & Deposing a tyrant. \\

  \end{xpchart}

}

\newcommand\joyXP{
  \begin{xpchart}{\Glsentrytext{joygod}}

    1 & Playing a prank. \\

    1 & Donating at least 1 \gls{gp} to the church. \\

    1 & Drinking and eating to excess. \\

    1 & Making a particularly bad pun. \\

    1 & Giving food or shelter to the needy. \\

    1 & Reminding another player about something in their Code to get them \glspl{xp}. \\

    3 & Winning a drinking competition. \\

    3 & Lifting the spirits of the downtrodden. \\

    3 & Creating a funny song. Requires at least a full night and an Intelligence + Performance action, \gls{tn} 12. \\

    3 & Playing a prank set up last session. \\

    5 & Hosting a feast for a village. \\

    5 & Creating a new type of alcohol. \\

    5 & Saving someone's life. \\

    10 & Saving a village or larger populated area from destruction. \\

    15 & Playing a prank set more than two sessions ago at the start of the campaign. \\

  \end{xpchart}


}

\newcommand\knowledgeXP{
  \begin{xpchart}{\Glsentrytext{knowledgegod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Learning a new secret. \\

    1 & Correcting someone on world-lore. \\

    1 & Gaining a new level in Academics or any sphere in the Path of Alchemy. \\

    1 & Remembering an \gls{npc}'s name. \\

    1 & Crafting a new magical item. \\

    1 & Drawing a map for the current session. \\

    1 & Overcoming a tricky situation. \\

    3 & Solving a complicated puzzle. \\

    5 & Donating at least 10 \gls{gp} to the temple. \\

    5 & Uncovering a conspiracy. \\

    5 & Writing an informative book on some topic. Intelligence + Academics is rolled at \gls{tn} 16 during downtime, and requires at least a month. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    10 & Finding and preserving important knowledge that would otherwise have been destroyed forever. \\

    15 & Solving a legendary puzzle. \\

  \end{xpchart}


}

\newcommand\experienceXP{
  \begin{xpchart}{Experience}

    1 & Finding a new type of food or drink. \\

    1 & Witnessing a flower open. \\

    1 & Seeing a change in Season. \\

    1 & Greeting a new type of humanoid. \\

    1 & Entering a new area, days away from anything seen before. \\

    1 & Hearing a new language. \\

    1 & Learning a new type of creative skill, such as Crafts or Performance. \\

    3 & Experiencing any emotion to heights never reached before. \\

    3 & Learning a Skill or sphere to a level above any other Skill or sphere you have. \\

    3 & Making a new friend. \\

    3 & Returning home after a longer period away than ever before. \\

    3 & Finding a mana lake. \\

    3 & Seeing a new magical item activate. \\

    5 & Experiencing deep tragedy. \\

    5 & Taking \gls{hp} Damage for the first time. \\

    5 & Falling in love. \\

    5 & Creating a masterpiece of some kind at \gls{tn} 14.
    Each Skill can only be used to gain \gls{xp} once. \\

    5 & Discovering a lost piece of art or literature. \\

    10 & Finding an area lost to all contact for more than a century. \\

    15 & Finding an area never before visited by people. \\

  \end{xpchart}
}

\newcommand\natureXP{
  \begin{xpchart}{\Glsentrytext{naturegod}}

    1 & Donating at least 1 \gls{cp} to the temple. \\

    1 & Hunting one's own food and dedicating it to \glsentrytext{naturegod}. \\

    1 & Gaining a new level in the Caving, Seafaring, or Wyldcrafting Skills. \\

    3 & Building a shrine -- requires 3 days work and an Intelligence + Crafts action, \gls{tn} 8. \\

    3 & Donating all of one's money to the temple. \\

    3 & Freeing a wild animal from captivity. \\

    3 & Destroying an `unnatural' creature such as the undead. \\

    5 & Composing a song to \gls{naturegod} -- requires an Intelligence + Performance action, \gls{tn} 15. \\

    5 & Seeing a new type of creature. \\

    10 & Establishing a new temple. \\

    15 & Saving some miles of land from being despoiled. \\

  \end{xpchart}


}

\newcommand\warXP{
  \begin{xpchart}{\Gls{wargod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Finding an interesting battle trophy. \\

    1 & Finishing a battle with 0 \gls{fp}. \\

    1 & Finishing a round with -10 \glspl{ap} or less. \\

    1 & Fighting a new type of creature. \\

    1 & Gaining a new level of the Combat Skill. \\

    1 & Surviving a skirmish while outnumbered. \\

    1 & Going first in the party when entering a dangerous situation. \\

    3 & Answering a one on one challenge. \\

    3 & Killing three opponents single handedly in one battle. \\

    3 & Killing a more dangerous opponent than ever before (danger is measured in \gls{xp}). \\

    3 & Surviving a large scale battle while outnumbered. \\

    5 & Donating at least 10 \gls{gp} to the temple. \\

    5 & Killing five or more opponents single handedly in battle. \\

    5 & Killing a dangerous opponent (16+ \gls{xp}) single handedly. \\

    10 & Killing a dangerous opponent (10+ \gls{xp}) without wearing armour. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    10 & Defeating a previously victorious opponent. \\

    15 & Starting a war. \\

  \end{xpchart}


}

\newcommand\deathXP{
  \begin{xpchart}{\Glsentrytext{deathgod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Cursing a dying enemy, to prevent them from achieving a proper afterlife. \\

    1 & Gaining a new level in the Necromancy sphere. \\

    1 & Giving someone blessings upon their death bed. \\

    1 & Losing \glspl{hp}. \\

    1 & Toasting to absent friends. \\

    3 & Performing an outlandish burial, with sacrifices and words appropriate for the deceased. \\

    3 & Easing the last moments of a dying person. \\

    3 & Erecting a shrine to the dead. Requires an Int + Craft action, \gls{tn} 9, at at least 10 gp. \\

    3 & Learning the history of a new area. \\

    3 & Protecting the graves of the honoured dead. \\

    5 & Donating at least 10 \gls{gp} to the temple of \gls{deathgod}. \\

    5 & Falling below 0 \gls{hp}. \\

    5 & Retrieving a dead body for a proper burial. \\

    10 & Building or funding a grand mausoleum. Takes a year and requires at least 100 gp. \\

    10 & Falling below -3 \glspl{hp}. \\

    15 & Falling below -5 \glspl{hp}. \\

  \end{xpchart}

}

\newcommand\tribeXP{
\begin{xpchart}{the Tribe}

  1 & Helping out a member of the tribe. \\

  1 & Greeting a member of the tribe during a long time away. \\

  1 & Welcoming a friend into the tribe. \\

  1 & Testing a potential new member of the tribe. \\

  1 & Defending your tribe's honour. \\

  3 & Spreading the honourable name of the tribe to outsiders. \\

  3 & Entering battle simply for the sake of the tribe. \\

  5 & Forming an alliance for the tribe. \\

  5 & Returning home after an extended trip away. \\

  5 & Acquiring a new home for your family. \\

  5 & Saving a member of the tribe from some terrible situation. \\

  5 & Fulfilling the final wishes of an ancestor.
  The ancestor's wish can be specified only while spending \glspl{storypoint}, and the \gls{xp} can only be collected if the mission is completed in a later Story. \\

  10 & Starting a family. \\

  15 & Saving one's entire tribe from destruction. \\

\end{xpchart}

}

\newcommand\exampleRandomCharacter{
  \setcounter{str}{1}
  \setcounter{dex}{0}
  \setcounter{spd}{0}
  \setcounter{int}{0}
  \setcounter{wts}{-1}
  \setcounter{cha}{1}

  \renewcommand\concept{Knightly Poet}
  \renewcommand\race{Human}
  \renewcommand\culture{Quennome}
  \renewcommand\name{Sean}
  \renewcommand\code{Experience}

  \setcounter{Academics}{1}
  \setcounter{Athletics}{0}
  \setcounter{Caving}{0}
  \setcounter{Crafts}{0}
  \setcounter{Deceit}{0}
  \setcounter{Empathy}{1}
  \setcounter{Medicine}{0}
  \setcounter{Performance}{1}
  \setcounter{Larceny}{0}
  \setcounter{Seafaring}{0}
  \setcounter{Stealth}{0}
  \setcounter{Tactics}{0}
  \setcounter{Vigilance}{0}
  \setcounter{Wyldcrafting}{0}
  \setcounter{Combat}{2}
  \setcounter{Projectiles}{0}

  \setcounter{fp}{5}

  \renewcommand\characterWeapons{\longsword}
  \renewcommand\characterArmour{Chain Mail & 4 & P & 1}
  \renewcommand\characterEquipment{\Gls{adventure} Equipment \Repeat{3}{{\Large\sqn} }}
  \setcounter{gold}{288}

  \settoggle{examplecharacter}{true}
  \settoggle{bestiarychapter}{true}
  \input{config/CS.tex}
  \settoggle{bestiarychapter}{false}
  \settoggle{examplecharacter}{false}
  \renewcommand\csComments{}
  \clearpage


}
