\begin{multicols}{2}

\subsection*{The Primordial Forest}

\begin{exampletext}

  The forest wants to eat you, so pay attention.

  You've had an easy life, in your secluded, little hole, but here we live above ground, where giant creatures crawl everywhere, and I can assure you I'm the gentlest one out here.
  They're probably looking at you and licking their lips.
Well, not `lips'.
  I don't think a single one has lips.
  Some have beaks, others have `mandibles', but you get my point.

  The chitincrawlers lay webs, but don't think they need to wait for you.
  If they get hungry enough, they'll run straight at you, stab at you with those pincers, and just start eating!

  And remember to be on the lookout for moving trees.
  If you see something shift that's meant to stay still, it could be a woodspy -- like an octopus, but\ldots do they have octopuses where you're from?
  No I didn't think so.
  But I'm glad to have you with us.
  I hear your people can see better in the dark than we can.

  Most of the world sits in darkness, just like this.
  Most of the world lacks roads, beer, beds, and everything that makes life worth living.
  For this reason, we exist, to push back the darkness, and make way for more civilization.

  \subsubsection*{Bandits}

  Notice the trees.
  \ifnum\value{temperature}<1
    They just sitting there in the \showSeason n wind right now, but once the warmer seasons hit, they'll be full of fruits, and you'll see things growing all around.
  \else
    There's good eating up there if you can climb.
    The forest is laying a trap for us, but it's a tasty one!
  \fi
  Humans could live out here like some kind of paradise, never working, just taking food from the trees -- at least over C\'alea and Laiquea, and the other warmer seasons.
  Even Toldea has plenty to eat if you know where to look!
  And all this means the forest has laid another trap for us.
  Thieves, cut-throats, and hags who want to escape the law come into the forest, and she treats them well.
  They live here, tax-free,
  \ifnum\value{temperature}>0
    and right now they're cold and desperate, and ready to watch where you sleep on the road.
  \else
    robbing villages over the cold seasons.
  \fi

  And obviously it's \emph{our} job to come out here and route them out, by fire and sword.


  \subsubsection*{And the fiends\ldots}

  Magic's a horrid thing.
  Once someone knows enough of it, they can destroy a city.
  And you can never spot someone who knows it.
  Well sometimes I think you can -- a shifty look in the eye, especially if they've been wandering past the \gls{edge} too long.
  They may as well stay out there, as far as I'm concerned.

  Sometimes we lose whole cities, to curses which bring beasts out of the forests, or maybe they'll just turn a town's walls to ice on a warm day and watch those walls melt while the beasts come in to eat the village.
  I've heard of dragons and giants, but I still say nothing spells trouble like a pissed-off hag.

  You don't know any magic do you?

  Shame.
  We could probably use some out here.
  At least a little blessing or something would be nice.
\end{exampletext}

\subsection*{The \Glsentrytext{edge}}

\begin{exampletext}
  We're coming up to civilization at last!
  See that patch of lawful land?
  No trees, vines, bushes, or anything?
  That means the town must be close.

  I was in a proper \textit{big burn} once.
  We covered the area in oils last Laiquea \ifnum\value{temperature}=3 on a day as warm as this one \else on a scorcher of a day \fi and it burnt so high I swear it reached \gls{ainumar} and made the gods stink for a week of woodsmoke!

  \subsubsection*{Outer Villages}

  You see how this village has a big wooden wall?
  That's just for basilisks.
  The rest of the forest will crawl, fly, or slither over the top -- they don't mind the wall.

  The great clear areas around the \gls{edge} give us a buffer.
  Anything that wants to skitter over here in the daylight gets filled with arrows.
  Of course that won't always kill a basilisk, but it can drive them off.

  We'll see some sheep inside, maybe even cows, but there's never much meat here.
  You can take them out grazing only a short way, where it's safe, and the sight of so many animals always tempts something out of the forest sooner or later.
  Mostly, meat comes from the inside, while the outer circles send back wood, or forest fruits, or anything from the fields outside their walls.

  I'm not like you.
  I signed up to the \gls{guard} by choice, and started in a place just like this.
  We all shot at the targets each day, but I hit more than the other kids, so I got to sit with the archers while the rest worked the fields.

  During daylight, we took turns sleeping and watching.
  When the beasties came running out of the forest, ready to snatch up a farmer from the field, we'd all stand and aim our bows together.
  It's not easy, hitting something that far off, and making sure you don't hit your own.
  I didn't always manage\ldots but he's mostly fine nowadays, so you do what you can, you know?

  Most humans on the \gls{edge} learn the bow, or at least how to use a crossbow.
  Anyone who doesn't puts everyone else in danger.
  I grew up in a village like this one -- and we had a lot of sleepless nights, telling each other stories of famous adventurers from back when that sort of thing was still legal.

  At night they come in too.
  Archers still stay up the top, and on still nights we listened out for the chimes.
  Villagers make them out of shells, sometimes old bones, or the hollowed-out head of a chitincrawler.
  Fill those things with rocks and they make a proper clack-clack!

  Of course you can't see anything at night, so you're just listening for the chimes, waiting to string the bow if one goes off.
  And mostly it's the wind, but you string the bow anyway, and wait for another chime, or the screams.
  The woodspies look in houses quietly, crawling around the rooves, feeling for unlatched doors or open windows, and then reach in, and try to pull someone out quickly.
  Of course every home has small windows, but sometimes they'll manage to crack the wooden frame open, or just grab a baby.
  Once the screaming starts, the creature gives you maybe one shot, maybe two, then it's off into the darkness.

  It turned out someone had left his window open, and a child next to it.
  Wasn't his first time, or first crime, and soon enough we voted on him.
  I didn't start it, but I could see everyone waiting for me to speak, so I said what they wanted.
  I said I was out there most night, with no sleep (just a small exaggeration), and he never did took his share of the work outside.
  And he should join me, take up the bow and whatnot.

  We didn't have any more armour for him, but it wouldn't have made much difference.
  I never saw what took him, but without any training, we basically fed him to the forest.
  
  Anyway, I guess that's why they call us the \gls{guard}.
  I felt glad when I joined the proper guild.
  Means people like you get a chance -- some training, maybe armour, and you go out with a lot of others the first few times.
  We can't stay long, so get some rations, and we'll be on the road soon, headed inwards.

  \subsubsection*{Lonely Roads}

  That's another one of your duties, recruit -- maintain the roads.
  Sometimes bandits slip past the outer villages and camp at the side.
  Sometimes critters in the forest do the same.
  A lot of them are smart enough to know where we go, so they'll sit at the side of the road, picking off traders who carry meat, or just any trader.
  Traders \emph{are} meat as far as the forest is concerned.

  I wanted to be a trader once, but honestly couldn't summon up the courage.
  I survived a couple of trips, but I knew I wouldn't make it for long.
  So if you're ever travelling out to the \gls{edge}, remember to let the word around town, just like I did there.
  The more people who travel together, the safer.
  And if you end up getting eaten by something, maybe the beast will leave the trader alone, and let him get to market.
  Then you'll die a hero!

  Everyone dies a hero in the \gls{guard}.

  See that crossroads ahead?
  That's a good sign.
  We passed the \gls{edge}, now we have two roads, meaning at least two villages around us.
  They'll have walls of their own too, but the farther inwards we get, the safer.

  Sometimes these outer roads break.
  At first you notice nobody is coming to visit the town from that direction.
  Then you notice that nobody who went that way returns\ldots
  A couple of weeks later, and people hold the wakes for anyone who went that way, and hopefully they have a different road out.

  %!
  When a village just has one road, then no roads, they just sit there like an island.
  Hopefully someone notices, and \glspl{guard} get to them; but until then they live on alone, without iron, pitch, or any other help from the outside.
  If that goes on too long, it's another win for the forest, and a loss for civilization.
  We'll keep on pushing in, but when we reach too far in, the forest eats our fingers.

  \subsubsection*{Quiet Hamlets}

  We're getting closer.
  See that little hamlet?
  No walls, or nothing -- just stone houses for emergencies.
  Very little makes it in this far.

  Whether it's beasts or bandits, they get tempted by the smells along the road, and end up in an altercation with one of the settlements further out.
  Even if something nasty made it in here, they eat the sheep before the people.
  Mostly.

  These inner lands provide most of the meat of Fenestra.
  I bet you've even had some back home.
  No?
  Well lets go up and say `hello'!
  Villagers always give hospitality to the \gls{guard} when they see us.

  \subsubsection*{Little Masters}

  Each area has its own wardens.
  It's not true what they say about humans -- we don't need leaders telling us what to do, but we have them anyway.
  They don't really do much, but I guess they look nice and fancy.
  Village wardens own a few villages, and town wardens own a town.

  Must be nice.
  Pointless, but nice.
\end{exampletext}

\subsection*{Hungry Towns}

\begin{exampletext}

  I suppose you've never seen a big city like Arthur's Wing before.
  No monsters live here, so everyone can rest easy, aside from the cutthroats and thieves, who of course have to worry about the likes of \emph{me} dragging them into our merry little crew and our glorious missions.
  Look at that pathetic beggar over there, asking for food.
  He can clearly walk, but refuses to sign up with us and fight for civilization.
  Even if he got eaten by something, it'll slow that something down while everyone else kills it or gets to safety.
  It's a good deed.
  He could be a hero.
  Everyone dies a hero in the \gls{guard}.

  \subsubsection*{Guilded Temples}

  I guess you know, when the forest eats you, Sylf collects your soul, and turns you into a monster.
  Well the barracks we're going to is also our temple.
  A \gls{guard} temple to ward \emph{her} off.

  Look over there -- that's an image of Sable, the god who takes your soul when you die of cold.
  Inside you can listen at the Temple of the Weavers -- the Guild that makes clothes to ward off Sable.
  Go tell them a story from your home if you can.
  They love hearing new stories, and you'll need their help when the deep cold sets in.
  Then later, we can see if a doula might bless your start in the \gls{guard}.

  All the temples in town want to save your soul, except maybe the Poison Guild, they just sell ale.
  Personally, I want to go with Eldren.
  If you die of old age, he takes you up to a peaceful land forever, and that's where I want to go.
  Just need to make sure none of the other gods claim me first.

  Let's get some rest.
  You've got a mission already.
  Nobody's come from Greenwell in a week, and someone needs to find out why.
  I've found a few other new recruits, so you won't be lonely.

  Time to be a hero.

\end{exampletext}

\end{multicols}
