\mapentry[trialStarts]{Through the blind tunnel}
you walk until the firelight, ahead.
The tall walls have a copper cage, curling over the top, to protect the `jury'.
The walls have four corridors, with metal gates, including yours.

Your gate falls as someone climbs the far cage, shouting ``oooooh!''.
The jester jingles as he climbs, shouting excitedly.

``This one's ahead of schedule'', he says to the \gls{warden}, who sits in judgement at the far end.
Then he turns to the crowd.

``You know I like it when they're eager.''

The crowd laughs.
Some clap.

``Now the excuses!
Go on.
Tell us excuses.
Something pretty and new.''

\begin{description}
  \item[Argue your accuser's guilt]
  by hollaring the simple facts, and deducing their knowledge implies their guilt.
  Roll $2D6$ and add \roll{Strength}{Academics} against \tn[8].
  \begin{description}
    \item[Roll over 8]
    and go to part \vref{accuserEnters}.
    \item[Roll under 8]
    and turn to part \vref{intoTheGuard}.
    \item[Roll 8 exactly]
    and you have time to pick another tactic.
  \end{description}
  \item[Tell the \gls{warden} what you think of him]
  in part \vref{crawler}.
  \item[If you have a dagger]
  you can try to throw it at the distant \gls{warden} in part \vref{throwDagger}.
  \begin{description}
    \item[Success]
    leads to part \vref{crowdWild}.
    \item[Failure or a tie]
    leads to \vref{intoTheGuard}.
  \end{description}
  \item[To try a joke]
  roll $2D6$ and add \roll{Charisma}{Empathy} against \tn[8].
  \begin{description}
    \item[Roll over 8]
    and go to part \vref{accuserEnters}.
    \item[Roll under 8]
    and turn to part \vref{intoTheGuard}.
    \item[Roll 8 exactly]
    and you realize jokes aren't your thing, and pick another option.
  \end{description}
  \item[Respond with stony silence]
  and turn now to part \vref{accuserEnters}.
\end{description}


\mapentry[argue]{Arguing your case}
quickly makes your voice horse, as you have to shout over the crowd's laughter every time the jester pulls a yawn (each one more dramatic than the last).
But you can show, quite clearly, that your accuser's knowledge means they are guilty, not you.
Then the \gls{warden} speaks.

``You have established that your guilt is in question, and the accuser's guilt is not.
Therefore, you will provide a trial for each other.''

The gate to your left opens.
\Pgls{sunGuard} leads him out -- \composeHumanName\ -- the man so foolish, he thought he could win something in the \gls{court}.
He steps out, confused, but confident as ever, longsword in hand.
The \gls{sunGuard} walks over to hand you a longsword.

Make \pgls{combat} roll: $2D6$ plus \roll{Dexterity}{Melee} at \tn[11].
Add another +2~Bonus for having a longsword.

\begin{description}
  \item[Roll under 11]
  and take \dmg{3}~Damage.
  You can spend \glspl{fp} to reduce the Damage, but after the \glspl{fp}, that blade parts your flesh.
  \item[Roll 11]
  and you can decide to deal and receive Damage, or neither.
  \item[Roll over 11]
  and you cut him, dealing \dmg{0}~Damage plus your Strength~Bonus, plus 2 for the longsword.
\end{description}

Once you deal 7 or more Damage, your accuser dies.
Repeat until victory or death.

If you survive the fight, turn to part \vref{intoTheGuard}.

\mapentry[throwDagger]{Stepping closer}
you protest your innocence so that nobody suspects the plan.
The \gls{court}'s high walls mean that every \gls{step} forward lets you see less of the high \gls{warden}.
But at twenty \glspl{step}, you still see the head, so you pull the dagger from your boots.

The crowd gasps.
Roll \roll{Dexterity}{Projectiles} at \tn[11].

\begin{description}
  \item[Roll under 11]
  and turn to part \vref{crawler}.
  \item[Roll 11]
  and turn to part \vref{thouArtJester}.
  \item[Roll over 11]
  and turn to part \vref{wardenDead}.
\end{description}

\mapentry[wardenDead]{Silence Spills}
as the \gls{warden} clutches his throat.
With no orders, the \gls{court} remains paralysed.

Then the \gls{warden} points at a gate, and collapses into the arms of his entourage.
